---
title: "Brackets"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA)
library(tidyverse)
```

## Simple example

```{r}
x <- list(c(1,2,3),c(4,5,6))
print(x)
```

X is a list with two elements: 

1. A vector of numbers with values 1, 2, and 3  
2. A vector of numbers with values 4, 5, and 6

```{r}
single.1 <- x[1]
single.2 <- x[2]
double.1 <- x[[1]]
double.2 <- x[[2]]

print(single.1)
print(single.2)

print(double.1)
print(double.2)
```

The major difference between the single bracket and double bracket is the type of object they return; in particular, when the object you are indexing (x in this case) is a list.

The single bracket will return a list (of size one) for the corresponding element. A double bracket will return a vector of the corresponding element. Note that single.1's output has an additional [[1]] above it, indicating that it is a list, while double.1's output does not have this.

Which bracket you use depends on what object type you need. For example, the mean() function can only take in a vector of numbers, rather than a list. 

```{r}
mean(single.1)
mean(double.1)
```

Furthermore, the object type also changes how indexing works. For example, if you want to get the first two observations (values 1 and 2), it is much easier to do so using double:

```{r}
single.1[[1]][1]
single.1[[1]][2]

double.1[1]
double.1[2]
```


## String processing from class

Looking at the string-processing code from class, we have:

```{r}
filename <- system.file("extdata/murders.csv", package = "dslabs")
lines    <- readLines(filename)
x <- str_split(lines, ",")[1:3]
x
```

As we can see, the object we are indexing (x) is a list. Since we want the actual values of the first index, rather than a list (of size 1) containing the values, we use the double bracket:

```{r}
col_names <- x[[1]]
col_names
```

Compared to a list (of size 1) containing the values:

```{r}
x[1]
```

However, when we are subsetting x to remove the first index (the -1 operator), we must use the single index since we are operating on the list level. 

```{r}
x[-1]
```







